---
title: "Vignette 1 -- PIC counting with 10X Cell Ranger output"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/Shared/Example_dataset1/')
```

### content 
This Vignette contains an example to use PIC counting on 10X Cell Ranger output. 

### required libraries 
Please make sure the following libraries are installed and loaded for the analysis. 
```{r eval=FALSE}
library('data.table')
library('GenomicRanges')
library('Matrix')
library('PICsnATAC')
```

### Input files

Input files for PIC counting includes:

  * cell barcodes with metadata (`singlecell.csv`)
  * list of peak regions (`peaks.bed`)
  * fragment files (`fragment.tsv.gz`)

Below are the step-by-step instructions for obtaining these input from 10X Cell Ranger output files. Example datasets can be downloaded from 10X Genomics website [here](https://www.10xgenomics.com/resources/datasets/5-k-peripheral-blood-mononuclear-cells-pbm-cs-from-a-healthy-donor-next-gem-v-1-1-1-1-standard-2-0-0).

### Cell Barcodes
If we want to keep the cells from Cell Ranger filtering scheme:
```{r eval=FALSE}
meta.data <- read.csv('atac_pbmc_5k_nextgem_singlecell.csv', header = T )
# -- in your own Cell Ranger output, the file name will be 'singlecell.csv'
meta.data_filtered <- meta.data[meta.data$is__cell_barcode == 1,]
cells <- meta.data_filtered$barcode
```

### Peak set
If we want to keep the set of peaks from Cell Ranger:
```{r eval=FALSE}
peaks <- data.table::fread('atac_pbmc_5k_nextgem_peaks.bed',header = F)
# -- in your own Cell Ranger output, the file name will be 'peaks.bed'
colnames(peaks) <- c('seqname', 'start', 'end')
peak_sets = GenomicRanges::makeGRangesFromDataFrame(peaks)
```
Note, the codes will be the same for MACS2 called peaks, just change the file name

### Run PIC and save output
```{r eval=FALSE}
fragment_tsv_gz_file_location <- 'atac_pbmc_5k_nextgem_fragments.tsv.gz'
pic_mat <- PIC_counting(
  cells = cells,
  fragment_tsv_gz_file_location = fragment_tsv_gz_file_location,
  peak_sets = peak_sets)
saveRDS(pic_mat, 'PIC_mat.rds')
```
Note if you have multiple matrices, please make sure you save them as different files 

### Reference
If you used PIC-snATAC counting in your analysis, please cite our manuscript as:

Miao Z, Kim J. Is single nucleus ATAC-seq accessibility a qualitative or quantitative measurement?. _bioRxiv_. 2022.

